On finiteness of structures produced deterministically by oritatami co-transcriptional folding \thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~S.}

### 2018-11-09 paper.tex
Finished simulation Thm, added some explanation before and after
Added Proposition 1 (proof is simple, so can be moved to appendix). Finished proof of Theorem about simulation. Changed statement from 2D Turing machines to algorithms, because otherwise we may need to add many unnecessary details about how the input is given on the 2D tape, how we traverse it, etc.

\begin{theorem}\label{thm:OS_to_2dTM}
	There exists a Turing machine with 2-dimensional tape that simulates any oritatami system $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ in time ??. 
\end{theorem}
\begin{proof}
	Take any step in the computation, up to which some $i \ge 0$ first beads of $w$ have been stabilized, with the last bead at a point $p$. 
	The number of all possible elongations of the current conformation by the next $\delta$-beads is $(6 \times 5^{\delta-1}) \times ((2^4)^{\delta-1} \times 2^5) \in 2^{O(\delta)}$. 
	We can check for each of these elongations whether its arity is at most $\alpha$ or not and whether it is $R$-valid or not in time $O(\delta)$. 
\end{proof}

>>>>

\begin{proposition}\label{prop:check_validity}
	For any rule set $R$, arity $\alpha$ and conformation $C = (P,w,H)$ it is possible to check whether $C$ is $R$-valid and whether $C\in \mathcal{C}_{\leq \alpha}$ in time $\mathcal{O}(|H|\cdot|w|\cdot|R|)$.
\end{proposition}
\begin{proof}
	To check whether $C$ is $R$-valid:
	\begin{enumerate}
		\item FOR each $(i,j)\in H$:
		\item \hspace{1cm} IF $(w[i],w[j])\notin R$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|w|\cdot|R|)$ time for any reasonable representation of $w$ and $R$, hence the whole process takes $\mathcal{O}(|H|\cdot |w|\cdot|R|)$ time.	
	To check the arity constraint $C\in \mathcal{C}_{\leq \alpha}$: 
	\begin{enumerate}
		\item FOR each $i\in \{1,\dots,|w|\}$:
		\item \hspace{1cm} IF $\mathrm{degree}(i)=|\{j | (i,j)\in H \}|>\alpha$ THEN answer NO and HALT
		\item answer YES and HALT
	\end{enumerate}	
	Checking the condition in 2. can be done in $\mathcal{O}(|H|)$ time for any reasonable representation of $H$, hence the whole process takes $\mathcal{O}(|w|\cdot|H|)$ time.
\end{proof}
\begin{theorem}\label{thm:OS_to_2dTM}
	There is an algorithm that simulates any oritatami system $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ in time $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$. 
\end{theorem}
\begin{proof}
	Take any step in the computation, up to which some $i \ge 0$ first beads of $w$ have been stabilized, with the last bead at a point $p$. 
	The number of all possible elongations of the current conformation by the next $\delta$-beads is $(6 \times 5^{\delta-1}) \times ((2^4)^{\delta-1} \times 2^5) \in 2^{O(\delta)}$. 
	By Proposition~\ref{prop:check_validity}, we can check for each of these elongations whether its arity is at most $\alpha$ or not and whether it is $R$-valid or not in time $\mathcal{O}((2^4)^{\delta-1}\cdot2^5\cdot \delta\cdot|R|)=2^{\mathcal{O}(\delta)}\cdot|R|$.  Therefore, the total running time is $2^{\mathcal{O}(\delta)}\cdot |R|\cdot|w|$.
\end{proof}

\begin{corollary}
	For fixed $\delta$ and $\alpha$, the class of problems solvable by oritatami systems $(\Sigma, R, \delta, \alpha, \sigma, w)$ is included in $\mathrm{DTIME}(n^3)$.
\end{corollary}
\begin{proof}
	The claim follows from Theorem~\ref{thm:OS_to_2dTM} and the fact that $|R|$ is implicitly bounded by $|w|^2$.
\end{proof}

%\begin{corollary}\label{polytranscript}
%	Let $k$ be a non-negative number. Let $\Sigma$, $H$, $\delta$, $\alpha$ be fixed and the transcript length bounded polynomially by the seed length, i.e., $|w|\in O(|\sigma|^k)$. Then, the language of seeds $\sigma$ accepted by an OS $(\Sigma, R, \delta, \alpha, \sigma, w)$ is in $\mathrm{DTIME}(|\sigma|^{2k})$.
%\end{corollary}

%If the transcript length is polynomially bounded by the seed length, than the accepted language is in $\mathrm{P}$.

Because of the time hierarchy theorems, we know that $\mathrm{P}\subsetneq \mathrm{EXP}$, so we can conclude that OS which cannot deterministically fold transcripts of length exponential in the length of the seed are not computationally universal.
